---
title: "BayesTrace V2"
format:
  dashboard:
    theme: flatly
    css: ./www/styles_quarto.css
    orientation: columns
params:
  file_index: file_index
  file_path: file_path
  chains_burnin: chains_burnin
  downsample: downsample
  shinyfy: shinyfy
---

```{r libraries, echo=F, include=F}
# Load required packages
library(bayestrace)
library(quarto)
library(scales)
library(tidyverse)
library(plotly)
library(reactable)
library(coda)
library(ape)
library(visNetwork)
library(ggraph)
library(igraph)
library(tidygraph)
library(ggtree)
library(scatterpie)
library(ggiraph)
library(htmlwidgets)
#library(phylocanvas)
library(ggiraph)
library(data.table)

# establish if running via shiny or locally
eval_shiny<-params$shinyfy==TRUE
eval_local<-!eval_shiny

# pass on parameters
file_index<-params$file_index
file_path<-params$file_path
chains_burnin<-params$chains_burnin
downsample<-params$downsample
```

```{r theming, include=FALSE}

# global themes
theme_set(theme_classic()+
            theme(
              # grid and axes
              axis.line=element_blank(),
              panel.grid.major.y = element_line(color = "grey95",size = 0.5),
              # facet strips
              strip.background = element_rect(color=NA, fill=NA),
              strip.text = element_text(hjust = 0, size=12))
)


# run colors
run_colors <- c(
  "1"   = "#727cf5",
  "2"    = "#0acf97",
  "3" = "#fa5c7c",
  "4"  = "#ffbc00",
  "5"   = "#2e4057",
  "6"   = "#8d96a3"
)

# add more colours
more_colors<-hcl.colors(n=10, palette = "Dynamic")
names(more_colors)<-as.character(7:16)

run_colors<-c(run_colors, more_colors)
```


```{r shinyinput, include=FALSE, eval=eval_shiny}
### USE THIS CHUNK IF RUNNING FROM SHINY

# Define file paths and names
log_path<-file_index %>%
  filter(filetype=="log") %>%
  select(filename, filepath) %>%
  deframe()

sched_path<-file_index %>%
  filter(filetype=="schedule") %>%
  select(filename, filepath) %>%
  deframe()

stones_path<-file_index %>%
  filter(filetype=="stones") %>%
  select(filename, filepath) %>%
  deframe()

tree_path <- file_index %>%
  filter(filetype=="tree") %>%
  pull(filepath)

traits_path<-file_index %>%
  filter(filename == read_bt_header(log_path[1])[[1]] %>%
                        filter(Options=="Data File Name") %>%
                        pull(2)
         ) %>%
  pull(filepath)


# Define EXTRA burnin (proportion of chain to remove as burnin. Default is 0 (keep entire chain))
burnin<-ifelse(is.numeric(chains_burnin), chains_burnin, 0)


# Define how many (random) iterations to keep for plotting (reduces load, default = 10,000)
if(!is.numeric(downsample)){
  downsample=downsample
} else{downsample=10000}

```

```{r localfilepaths, include=FALSE, eval=eval_local}
### USE THIS BLOCK IF RUNNING LOCALY

### TEST DATASET
#file_path="~/Downloads/rjhp_cov_6states/"
#chains_burnin=0
#downsample=5000

# Define file paths and names
all_files<-list.files(path=file_path, full.names = T, recursive = T)

log_path<-all_files[grep(".Log.txt", all_files)]
names(log_path)<-basename(log_path) 
 

sched_path<-all_files[grep(".Schedule.txt", all_files)]
names(sched_path)<-basename(sched_path)

stones_path<-all_files[grep(".Stones.txt", all_files)]
names(stones_path)<-basename(stones_path)


tr_name<-read_bt_header(log_path[1]) %>%
                        filter(Options=="Tree File Name") %>%
                        pull(2)

tree_path<-all_files[grep(tr_name, all_files)][1] # keep only first tree file it finds

tra_name<-read_bt_header(log_path[1]) %>%
                        filter(Options=="Data File Name") %>%
                        pull(2)

traits_path <- all_files[grep(tra_name, all_files)][1] # keep only first traits file it finds

# make sure all exist
#file.exists(c(traits_path,tree_path,log_path,sched_path,stones_path))


```

```{r message=FALSE, warning=FALSE}
# load BayesTraits log file
bt<-read_bt_log(log_path = log_path)

# load input tree
tre<-read.nexus(file=tree_path)

## if multiPhylo, get majority rule consensus tree
if(class(tre)=="multiPhylo"){
  tre<-consensus(tre, p = 0.5)
}

# define run mode
run_mode<-case_when(
  any(str_detect(bt$header, "Discrete")) &  any(str_detect(bt$header, "MultiState")) ~ "Different run types loaded",
  any(str_detect(bt$header, "Discrete")) ~ "Discrete",
  any(str_detect(bt$header, "MultiState")) ~ "MultiState"
  )


# load traits file
if(run_mode=="MultiState") {
  dat<-read_tsv(file = traits_path,
              col_names = c("Species","Trait")
              ) %>%
      mutate(Trait=str_replace(Trait, "-", NA_character_))
}
if(run_mode=="Discrete") {
  dat<-read_tsv(file = traits_path,
              col_names = c("Species","Trait1","Trait2")
              ) %>%
    mutate(Trait1=str_replace(Trait1, "-", NA_character_),
           Trait2=str_replace(Trait1, "-", NA_character_),
           Trait=ifelse(is.na(Trait1)|is.na(Trait2), NA, paste(Trait1, Trait2, sep=",")))
}
 
# load schedule files
acceptance_rate<-read_bt_schedule(file_path = sched_path)

# load stones files
if(length(stones_path)>0){
  stones<-read_bt_stones(file_path = stones_path)
}
```

```{r}
# remove burnin and downsample 
bt$chain_clean<-bt$chain %>%
  group_by(`Run ID`) %>%
  slice_tail(prop = 1-chains_burnin) %>%
  ungroup() %>%
  filter(row_number() %in% floor(seq(1, n(), length.out=downsample)))

acceptance_rate$schedule<-acceptance_rate$schedule  %>%
  group_by(`Run ID`) %>%
  slice_tail(prop = 1-chains_burnin) %>%
  ungroup() %>%
  filter(row_number() %in% floor(seq(1, n(), length.out=downsample))) 

```


```{r}
# define colour space for descrete runs
if(run_mode=="Discrete"){
  traits_colors=tibble(trait=c("0,0","0,1","1,0","1,1"),
                      cols=run_colors[1:4]) %>%
    deframe()
} else{
  traits_colors<-run_colors[1:length(unique(na.omit(dat$Trait)))]
  names(traits_colors)<-unique(na.omit(dat$Trait))
}
```

```{r}
# define "mode". I.e. are multiple runs the same, or are they run with different settings 
comparison_mode<-bt$header %>%
  filter(if_any(2:ncol(.), ~ .x != bt$header[,2])) %>%
  filter(!Options %in% c("Log File Name","Seed", "Schedule File","Iterations","Cores")) %>%
  nrow()>0
```



# {.sidebar}

### Dashboards
[Data Input](#data-input)\
[MCMC Trace](#mcmc-trace)\
[MCMC Diagnostics](#mcmc-diagnostics)\
[Root State](#root-state)\
[Transition Rates](#transition-rates)\
[Node Reconstructions](#node-reconstructions)\
[Reverse Jump](#reverse-jump)\
[Model Comparison](#model-comparison)\

# Data Input

## {.tabset}

### Header

```{r input-header}
# cat table, highlighting differences
bt$header %>%
  mutate(not_same=if_any(2:ncol(.), ~ .x != bt$header[[2]])) %>% # identify columns that are not the same
  reactable(
    defaultPageSize = 50,
    defaultColDef = colDef(
      style = JS(
        "function(cellInfo) {
          if (cellInfo.row.not_same) {
            return {color: 'red'}
          } else {
            return {color: 'black'}
          }
        }"
      )
    ),
    columns = list(
      not_same = colDef(show=FALSE)
    )
  )
```

### Priors & Restrictions

::: {.card title="Prior"}
```{r input-prior}
bt$priors %>%
  reactable(defaultPageSize = 20)
```
:::


::: {.card title="Restrictions"}
```{r input-restrictions}
bt$restrictions %>%
  reactable(defaultPageSize = 20)
```
:::

### Reconstructions & Fossilizations

```{r input-reconstructions}
if(is.null(bt$tags) | is.null(bt$recon)) {
  cat("No tags or reconstructions set")
}else{
bt$recon %>%
  left_join(bt$tags) %>%
  reactable(defaultPageSize = 20,
            searchable = T)
  }
```

### Tree

```{r input-tree}
bt$tree %>%
  reactable(defaultPageSize = 20)
```

# MCMC Trace

::: {.card title="Chain Iterations" width="60%"}
```{r chain-plot}
gg_chain<-bt$chain_clean %>%
  ggplot(aes(x=Iteration, y=Lh, color=`Run ID`)) +
  geom_line(alpha=0.5) +
  geom_smooth(se = FALSE) +
  scale_colour_manual(values=unname(run_colors)) +
  theme(legend.position = "none")

# plotly-fy
ggplotly(gg_chain) %>%
  layout(legend = list(orientation = "h", x = 0, y =-0.2))
```
:::

::: {.card title="Chain Distributions" width="40%"}
```{r chain-violins}
gg_chain_violins<-bt$chain_clean %>%
  ggplot(aes(x=`Run ID`, y=Lh, fill=`Run ID`, alpha=0.5)) +
  geom_violin(trim = FALSE, aes(text=`Run ID`), color=NA) +
  geom_boxplot(width = 0.1, aes(color=`Run ID`)) +
  scale_fill_manual(values=unname(run_colors)) +
  scale_color_manual(values=unname(run_colors)) +
  labs(x="", y="Log Likelihood") +
  theme(legend.position = "none",
        axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())

## plotly-fy
ggplotly(gg_chain_violins, tooltip = "text") %>%
  layout(boxgap = 0.9)
  
```
:::

# MCMC Diagnostics

## Column  {width="50%"}

::: {.card title="Effective Sampling Size"}
```{r ees-table}
ESS<-bt$chain %>%
  select(`Run ID`,starts_with(c("Lh","q","Root"))) %>%
  group_by(`Run ID`) %>%
  summarize_all(.funs=function(x)ifelse(any(is.na(x)),NA,effectiveSize(x))) %>%
  ungroup() %>%
  pivot_longer(-`Run ID`, names_to = "Parameter", values_to="ESS") %>%
  pivot_wider(names_from = `Run ID`, values_from = ESS)
  
# plot
ESS %>%
  mutate_if(is.numeric, round, 2) %>%
  reactable(
    defaultPageSize = 50,
    defaultColDef = colDef(
    style = function(value) {
      if (is.character(value)) {
        color <- "black"
      } else if (!is.na(value) & value > 2000 ) {
        color <- "#008000"
      } else if (!is.na(value) & value < 2000) {
        color <- "#e00000"
      } else {
        color <- "black"
      }
      list(color = color)
    }
    )
  )
```
:::

::: {.card title="Mean Acceptance Rates"}
```{r mean-acceptance-table}
# mean values
acceptance_rate$schedule %>%
  group_by(`Run ID`) %>%
  select(`Rates - Accepted`, `Sample Ave Acceptance`, `Total Ave Acceptance`) %>%
  summarise_all(mean) %>%
  mutate_if(is.numeric, round, 2) %>%
  reactable()
```
:::

::: {.card title="Total Average Acceptance Rates"}
```{r acceptance-plot}
# plot of total average acceptances
gg_acc<-acceptance_rate$schedule %>%
  group_by(`Run ID`) %>%
  mutate(Iterations = 1:n(),
         `Total Ave Acceptance`=`Total Ave Acceptance`*100) %>%
  ggplot(aes(x=Iterations,y=`Total Ave Acceptance`, color=`Run ID`)) +
  scale_colour_manual(values=unname(run_colors)) +
  labs(y="Acceptance (%)") +
  geom_line() +
  theme(legend.position = "none")
  
# plotly-fy
ggplotly(gg_acc)
```
:::

## Column {width="50%"}

::: {.card title="ML Autocorrelation"}
```{r autocorr-plot}
# calculate auto correlations
bt_acf<-list()
for(i in unique(bt$chain_clean$`Run ID`)){
  tmp_acf<-acf(plot=FALSE,
               bt$chain_clean %>%
                 filter(`Run ID`==i) %>%
                 pull(Lh))
  bt_acf[[i]]<-tibble(Lag=c(tmp_acf$lag), acf=c(tmp_acf$acf))
}

# plot
gg_autocorr<-bind_rows(bt_acf,.id = "Run ID") %>%
  ggplot(aes(x=Lag, y=acf, color=`Run ID`)) +
  geom_hline(yintercept = 0, linewidth=0.5) +
  geom_point(size = 2) +
  geom_segment(aes(xend = Lag, yend = 0), size = 0.75) +
  scale_colour_manual(values=unname(run_colors)) +
  labs(y="Autocorrelation Function") +
  facet_wrap(~`Run ID`, ncol = 1) +
  #facet_grid(rows = vars(`Run ID`), switch = "y") +
  theme(strip.text.x = element_blank(),
        legend.position = "none")
  
#plotlyfy
ggplotly(gg_autocorr)
```
:::


# Root State

::: {.card title="Root State"}
```{r fig.cap='Root state probabilities for each run'}

if(comparison_mode) {
  gg_root_state<-
    bt$chain_clean %>%
    group_by(`Run ID`) %>%
    select(`Run ID`, starts_with("Root")) %>%
    summarise_all(mean) %>%
    pivot_longer(-`Run ID`,names_to="State",values_to="prob") %>%
    mutate(State=str_remove_all(State, ".*P\\(|\\)")) %>%
    ggplot(aes(y=`Run ID`,x=prob, fill=State)) +
    #geom_bar(stat="identity") +
   ggiraph:: geom_col_interactive(aes(tooltip = paste0("Run ID: ", `Run ID`,
                                            "</br>State: ", State,
                                            "</br>Prob: ", prob)))+
    coord_polar() +
    scale_fill_manual(values = traits_colors) +
    facet_wrap(~`Run ID`) +
    theme_void() + 
    theme(legend.position = "bottom") 
} else {
  gg_root_state<-
    bt$chain_clean %>%
     select(starts_with("Root")) %>%
     summarise_all(mean) %>%
     pivot_longer(everything(),names_to="State",values_to="prob") %>%
     mutate(State=str_remove_all(State, ".*P\\(|\\)")) %>%
     ggplot(aes(x="",y=prob, fill=State)) +
     #geom_bar(stat="identity") +
     ggiraph::geom_col_interactive(aes(tooltip = paste0("State: ", State,
                                            "</br>Prob: ", prob)))+
     coord_polar("y") +
     scale_fill_manual(values = traits_colors) +
     theme_void() + 
     theme(legend.position = "right")
}

# plot
#girafe(ggobj=gg_root_state, width_svg = 50, height_svg = 50)
gg_root_state
```
:::


# Transition Rates

## {.tabset}

### Summarized Transition Rates

::: {.card title="Transition Rates Network" height="70%"}
```{r trans-network}
# calculate mean transition rates based on all concatenated chains if the runs are the same, if not, then transitions are shown for each chain separately

if(!comparison_mode){
  
 transition_dat<-bt$chain_clean %>%
  select(starts_with("q")) %>%
  summarise_all(mean) %>%
  pivot_longer(everything(),names_to="parameter", values_to = "rate") %>%
  mutate(from=str_sub(parameter, 2,2), to=str_sub(parameter, 3,3)) %>%
  arrange(from)

transition_nodes<-tibble(
  id = unique(c(transition_dat$from, transition_dat$to)),
  label = id,
  Run_ID = id,
  color = traits_colors[id],
  shape="circle",
  font.color="white"
)

transition_edges<-transition_dat %>%
  select(from, to, rate) %>%
  mutate(arrows="to",
         label=as.character(round(rate,3)),
         width=scales::rescale(rate, to = c(1, 5)))

visNetwork::visNetwork(transition_nodes, transition_edges) %>%
  visPhysics(solver = "forceAtlas2Based", 
             forceAtlas2Based = list(gravitationalConstant = -200)) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visOptions(selectedBy = list(variable="id", multiple=T))


} else {
  
  transition_dat<-bt$chain_clean %>%
    group_by(`Run ID`) %>%
    select(starts_with("q")) %>%
    summarise_all(mean) %>%
    pivot_longer(-`Run ID`,names_to="parameter", values_to = "rate") %>%
    mutate(from=str_sub(parameter, 2,2), to=str_sub(parameter, 3,3)) %>%
    arrange(from)

  transition_nodes<-transition_dat %>%
    select(`Run ID`, from, to) %>%
    pivot_longer(-`Run ID`, values_to="label") %>%
    select(-name) %>%
    distinct_all() %>%
    rename(Run_ID=`Run ID`) %>%
    mutate(id=paste0(Run_ID, label),
           color=traits_colors[label],
           shape="circle",
           font.color="white")
    
  if(run_mode=="Discrete") {
    
    transition_nodes<-transition_nodes %>%
      mutate(label=case_when(
        label=="1" ~ "0,0",
        label=="2" ~ "0,1",
        label=="3" ~ "1,0",
        label=="4" ~ "1,1"),
        color=traits_colors[label])
  }

transition_edges<-transition_dat %>%
  select(`Run ID`, from, to, rate) %>%
  mutate(arrows="to",
         from=paste0(`Run ID`,from),
         to=paste0(`Run ID`, to),
         label=as.character(round(rate,3)),
         width=scales::rescale(rate, to = c(1, 5)))

visNetwork::visNetwork(transition_nodes, transition_edges) %>%
  visPhysics(solver = "forceAtlas2Based", 
             forceAtlas2Based = list(gravitationalConstant = -100)) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visOptions(selectedBy = list(variable="Run_ID", multiple=T))
}

```
:::

::: {.card title="Transition Rates Matrix" height="30%"}
```{r trans-table}
bt$chain_clean %>%
  select(starts_with("q")) %>%
  summarise_all(mean) %>%
  pivot_longer(everything(),names_to="parameter", values_to = "rate") %>%
  mutate(from=str_sub(parameter, 2,2), to=str_sub(parameter, 3,3)) %>%
  arrange(to) %>%
  select(from, to, rate) %>%
  pivot_wider(names_from = to, values_from = rate) %>%
  arrange(from) %>%
  mutate_if(is.numeric, round, 2) %>%
  reactable::reactable()
```
:::


### Transition Rate Densities

::: {.card title="Transition Rates Densities"}
```{r trans-densities}

if(comparison_mode) {
  rate_density_gg<-bt$chain_clean %>%
    group_by(`Run ID`) %>%
    select(starts_with("q")) %>%
    pivot_longer(-`Run ID`, names_to = "transition", values_to = "rate") %>%
    ggplot(aes(x=rate, fill=`Run ID`)) +
    geom_density(alpha=0.25, color=NA) +
    geom_vline(data=. %>%
               group_by(`Run ID`, transition) %>%
               summarise(rate=mean(rate)),
             aes(xintercept=rate, linetype="dashed", colour=`Run ID`)) +
    scale_fill_manual(values=unname(run_colors)) +
    scale_color_manual(values=unname(run_colors)) +
    facet_wrap(~transition, scales = "free") +
    theme(legend.position = "none")
  
  #plotlyfy
  ggplotly(rate_density_gg)
} else {
  
  rate_density_gg<-bt$chain_clean %>%
  select(starts_with("q")) %>%
  pivot_longer(everything(), names_to = "transition", values_to = "rate") %>%
  ggplot(aes(x=rate)) +
  geom_density(alpha=0.75, fill="grey60",color=NA) +
  geom_vline(data=. %>%
               group_by(transition) %>%
    summarise(rate=mean(rate)),
    aes(xintercept=rate, linetype="dashed")) +
  facet_wrap(~transition, scales = "free") +
  theme(legend.position = "none")
  
  #plotlyfy
  ggplotly(rate_density_gg)
}

```
:::

# Node Reconstructions

```{r calc-anc}
## add trait data to tree
tre$traits<-dat %>%
  arrange(match(Species, tre$tip.label)) %>%
  pull(Trait)

## conditionally add node states

if(is.null(bt$tags)) {
  cat("No nodes were set for ancestral state reconstrucitions")
}else{
  
# make the ancestral states df
anc_df<-bt$tags %>%
  left_join(bt$recon, by="Tag")


# get node numbers for ancestral states
node_num<-c()
for(i in 1:nrow(anc_df)){
  tips<-unlist(str_split(anc_df$Species[i], " "))
  node_num[i]<-getMRCA(tre, tips)
}

## add node numbers and extract mean probabilities per node state
anc_df<-anc_df %>%
  mutate(node_num=node_num) %>%
  left_join(
    bt$chain %>%
      select(contains(anc_df$Node)) %>%
      summarise_all(mean, na.rm=TRUE) %>%
      pivot_longer(everything()) %>%
      mutate(name=str_remove_all(name, "P\\(|\\)")) %>%
      separate(name, into=c("Node","Prob"), sep=" ") %>%
      pivot_wider(values_from = value, names_from = Prob),
    by="Node")

}
```


## {.tabset}

### Tree
```{r anc-tree}

big_tree<-length(tre$tip.label)>100

plot(tre,
     show.tip.label = !big_tree,
     label.offset = ifelse(big_tree, 0.1, 0.5),
     no.margin = T,
     type=ifelse(big_tree,"fan","phylogram"),
     edge.width = ifelse(big_tree, 0.25, 1)
     #tip.color = traits_colors[tre$traits]
)
tiplabels(pch=16, col=traits_colors[as.character(tre$traits)], cex=ifelse(big_tree, 0.25, 3))

if(!is.null(bt$tags)){
  nodelabels(node = anc_df$node_num,pie = anc_df %>%
               select(any_of(na.omit(unique(dat$Trait)))),
             cex = 1,
             piecol=traits_colors)
  legend("bottomleft", legend = names(traits_colors), pch=15, col=traits_colors, bty="n")
}
```

### Table
```{r anc-table}
if(exists("anc_df")){
  anc_df %>%
    select(Tag, Node, Species,any_of(unique(dat$Trait) %>% na.omit())) %>%
    mutate_if(is.numeric, round, 2) %>%
    reactable(searchable = T)
} else {
  cat("No nodes were set for ancestral state reconstrucitions")
}
```


# Reverse Jump

```{r check-rj}
## check if RJ has been used
rj<-"Model string" %in% colnames(bt$chain_clean)
```


## Column  {width="50%"}

::: {.card title="Rev Jump Rates Sampling"}
```{r rj-sampling}
if(rj){
transition_rates<-bt$chain_clean %>%
  select(starts_with("q")) %>%
  pivot_longer(everything(), names_to = "transitions", values_to = "rates") %>%
  group_by(transitions) %>%
  summarise(`Mean Rate`=mean(rates),
            `Median Rate`=median(rates),
            `% Non-Zero`=(sum(rates!=0)/length(rates))*100,
            ESS=effectiveSize(coda::mcmc(rates)),
            HPD=HPDinterval(coda::mcmc(rates)) %>%
              as_tibble() %>%
              mutate_all(round, 2) %>%
              unite("x",c(lower, upper),sep="-") %>% pull(x)
            ) %>%
  arrange(desc(`% Non-Zero`)) %>%
  mutate_if(is.numeric, round, 2)

# plot
transition_rates %>%
  reactable()
} else {
  cat("No Reverse Jump used")
}
```
:::

::: {.card title="Rev Jump Model Ranking - Table"}
```{r rj-ranking-table}
if(rj){
  model_ranking<-bt$chain_clean %>%
    mutate(`Model string`=str_remove(`Model string`, "'")) %>%
    count(`Model string`, name = "N sampled") %>%
    mutate(`% sampled`=`N sampled`/sum(`N sampled`)*100) %>%
    arrange(desc(`N sampled`)) %>%
    mutate(ranking=row_number()) %>%
  mutate_if(is.numeric, round, 2) 

# plot table
model_ranking %>% 
  reactable()
}else {
  cat("No Reverse Jump used")
}
```
:::

::: {.card title="Rev Jump Model Ranking - Barchart"}
```{r rj-ranking-plot}
if(rj){
  revj_ranking_gg<-bt$chain_clean %>%
    count(`Model string`, name = "N sampled") %>%
    mutate(`% sampled`=`N sampled`/sum(`N sampled`)*100) %>%
    arrange(desc(`N sampled`)) %>%
    slice_head(n=10) %>%
    mutate(`Model Ranking`=as.factor(row_number()),
           `Model string` = str_remove(`Model string`, "'")) %>%
    ggplot(aes(x=`Model Ranking`, y=`% sampled`,
               text=paste0("<br>Model Ranking: ", `Model Ranking`,
                           "<br>Model String: ", `Model string`,
                           "<br>% Sampled: ", `% sampled`)
               )) +
    geom_bar(stat="identity")
  
  #plotlyfy
  ggplotly(revj_ranking_gg, tooltip = "text")
}else {
  cat("No Reverse Jump used")
}
```
:::

## Column {width="50%"}

::: {.card title="Non-Zero Transitions"}
```{r rj-network-nonzero}

if(rj){
  ### Plot transition network for 0% non-zero 

  transition_dat<-transition_rates %>%
  mutate(from=str_sub(transitions, 2,2),
         to=str_sub(transitions, 3,3)) 
  
transition_nodes<-tibble(
  id= unique(c(transition_dat$from, transition_dat$to)),
  label=id,
  group=id,
  color=traits_colors[id],
  shape="circle",
  font.color="white"
)

# make colour ramp
rbPal <- colorRampPalette(c('skyblue','#301e5d'))

transition_edges<-transition_dat %>%
  select(from, to, `% Non-Zero`) %>%
  mutate(id=row_number(),
         arrows="to",
         color= rbPal(nrow(.))[as.numeric(cut(`% Non-Zero`,breaks = nrow(.)))],
         label=as.character(round(`% Non-Zero`,2)),
         width=scales::rescale(`% Non-Zero`, to = c(1, 5)))

visNetwork(transition_nodes, transition_edges) %>%
  visPhysics(solver = "forceAtlas2Based", 
             forceAtlas2Based = list(gravitationalConstant = -100)) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visOptions(selectedBy = list(variable="id", multiple=T))


} else {
  cat("No Reverse Jump used")
}

```
:::


::: {.card title="Top 3 Most Sampled Models"}
```{r rj-network-top3}
if(rj){

  transition_dat<-bt$chain_clean %>%
    select(`Model string`,starts_with("q")) %>%
    group_by(`Model string`) %>%
    summarise_all(mean) %>%
    pivot_longer(-`Model string`, names_to="parameter", values_to = "rate") %>%
    mutate(from=str_sub(parameter, 2,2),
           to=str_sub(parameter, 3,3),
           `Model string`=str_remove(`Model string`, "'")) %>%
    left_join(model_ranking) %>%
    # add transition category
    group_by(`Model string`) %>%
    mutate(rate_category=str_split(`Model string`[[1]], " ")[[1]][row_number()]) %>%
    ungroup() %>%
    ## arrange by ranking and keep only top 3
    arrange(ranking, from) %>%
    filter(ranking %in% c(1:3))
  
  # make nodes table
  transition_nodes<-transition_dat %>%
    select(`Model string`, from, to, ranking) %>%
    pivot_longer(-c(`Model string`, ranking), values_to="label") %>%
    select(-name) %>%
    distinct_all() %>%
    rename(group=`Model string`,
           "Model_Ranking" = ranking) %>%
    mutate(id=paste0(group, label),
           color=traits_colors[label],
           shape="circle",
           font.color="white")
  

transition_edges<-transition_dat %>%
  select(`Model string`, from, to, rate, rate_category) %>%
  filter(rate!=0) %>%
  mutate(id=row_number(),
         arrows="to",
         from=paste0(`Model string`,from),
         to=paste0(`Model string`, to),
         label=as.character(round(rate,3)),
         color=rate_category,
         width=scales::rescale(rate, to = c(1, 5))) 

visNetwork::visNetwork(transition_nodes, transition_edges) %>%
  visPhysics(solver = "forceAtlas2Based", 
             forceAtlas2Based = list(gravitationalConstant = -100)) %>%
  visInteraction(navigationButtons = TRUE) %>%
  visOptions(selectedBy = list(variable="Model_Ranking", multiple=T)) 
  
} else {
  cat("No Reverse Jump used")
}
par(mfrow=c(1,1))

```
:::

# Model Comparison


```{r mc-check}
# check if stones file exists
st<-exists("stones")
multi_stones<- length(stones_path)>1
```

## Column  {width="50%"}

::: {.card title="Stepping Stone Sampler" height="50%"}
```{r}
if(st){
  stones$header %>%
    reactable()
} else {
  cat("No stepping stones sampler set")
}
```
:::

::: {.card title="Log Marginal Likelihood"}
```{r}
if(st){
  stones$`marginal likelihood` %>%
    mutate_if(is.numeric, round, 2) %>%
    reactable(defaultPageSize = 2)
} else {
  cat("No stepping stones sampler set")
}
```
:::

::: {.card title="Log Bayes Factor"}
```{r}
if(st & multi_stones){
  
## calculate log Bayes Factor (x1=complex model, x2=simple model)
LogBF<-function(mod1,mod2){2*(mod1-mod2)}

### rank models in decreasing complexity 

if(run_mode=="Discrete"){
mods_tmp<-bt$header %>%
    filter(Options=="Model") %>%
    select(-Options) %>%
    pivot_longer(everything()) %>%
    deframe() %>%
  sort()

mod1=names(mods_tmp)[str_detect(mods_tmp, "Discrete: Dependent")]
mod2=names(mods_tmp)[str_detect(mods_tmp, "Discrete: Independent")]

  
} else {
  ## if not discrete or otherwise specified, rank based on number of columns in the chains
  model_complexity<- bt$chain_clean %>%
  split(f = as.factor(.$`Run ID`)) %>%
  lapply(function(x) x[,colSums(is.na(x))<nrow(x)]) %>%
  lapply(function(x) x[,colSums(x != 0) > 0]) %>%
  sapply(.,ncol)

  ### select most complex and next most complex model
mod1=names(which.max(model_complexity))
mod2=names(which.min(model_complexity)) 

}


### calculate logBF (based on means if more than one likelihood is present)
log_bf<-LogBF(mod1=stones$`marginal likelihood` %>%
                select(any_of(mod1)) %>%
                pull() %>%
                mean(na.rm=T),
              mod2=stones$`marginal likelihood` %>%
                select(any_of(mod2)) %>%
                pull() %>%
                mean(na.rm=T)
              ) %>%
  round(2)

# return Bayes Factor
cat(paste0("Log Bayes Factor: ",log_bf,
            "\nThe complex model (",mod1, ") is a ",ifelse(log_bf>0,"better","worse")," fit.")
            )
} else {
  cat("This needs at least two stones samplers")
}
```
:::

## Column {width="50%"}

::: {.card title="Stones Power"}
```{r fig.width=12}
if(st){
  stones_pow_gg<-stones$stones %>%
    ggplot(aes(x=`Stone No`, y=Power, colour=`Run ID`)) +
    geom_line() +
    scale_colour_manual(values=unname(run_colors)) +
    theme(legend.position = "none")
  
  #plotlyfy
  ggplotly(stones_pow_gg)
  
} else {
  cat("No stepping stones sampler set")
}
```
:::


::: {.card title="Stones Marginal Likelihood"}
```{r fig.width=12}
if(st){
  stones_marml_gg<-stones$stones %>%
    ggplot(aes(x=`Stone No`, y=`Stone MLh`, colour=`Run ID`)) +
    geom_line() +
      scale_colour_manual(values=unname(run_colors)) +
    theme(legend.position = "bottom") +
    theme(legend.position = "none")
  
  #plotlyfy
  ggplotly(stones_marml_gg)
  
} else {
  cat("No stepping stones sampler set")
}

```
:::


::: {.card title="Running Marginal Likelihood"}
```{r fig.width=12}
if(st){
  running_marml_gg<-stones$stones %>%
    ggplot(aes(x=`Stone No`, y=`Running MLh`, colour=`Run ID`)) +
    geom_line() +
      scale_colour_manual(values=unname(run_colors)) +
    theme(legend.position = "bottom") +
    theme(legend.position = "none")
  
  #plotlyfy
  ggplotly(running_marml_gg)
  
} else {
  cat("No stepping stones sampler set")
}

```
:::

